\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-27 tex/paper.tex
save

As illustrative experiments, we evaluate in this section adversarial variational
optimization on well-known distributions for continuous and discrete data.

We first consider inference for a univariate Gaussian with unknown mean and
unknown variance. The observed data is sampled from a univariate Gaussian with
mean $\mu = 1$ and variance $\sigma^2 = 0.5^2$. Algorithm~\ref{alg:avo} is run
for 300 epochs with mini-batches of size $M=64$ and the following configuration.
For the critic, we use a 3-layer MLP with 10 hidden nodes per layer and ReLU
activations. At each epoch, Adam is run for $n_\text{critic}=100$ iterations
with a step size $\alpha=0.01$, decay rates $\beta_1=\beta_2=0.5$ and its inner
first and second moment vectors reset at each epoch in order to avoid building
momentum in staled directions.  For modeling $\mu$ and $\sigma^2$, we use
univariate Gaussian proposal distributions $q(\mu|\psi)$ and $q(\log
\sigma|\psi)$ initialized with zero mean and unit variance. At each epoch,
parameters $\psi$ are updated by making one Adam step, with $\alpha=0.01$ and
$\beta_1=\beta_2=0.1$. Penalty coefficients are set to $\lambda=0.025$ and
$\gamma=5$.

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/gaussian-gamma=5.pdf}
    \caption{\glnote{todo}}\label{fig:gaussian}
\end{figure}

@@

Similarly, ... \glnote{discrete case with a Poisson distribution.}

>>>>

As illustrative experiments, we evaluate adversarial variational
optimization on well-known distributions for continuous and discrete data.
We consider these distributions as parameterized simulators, from which
we can only generate data.

For continuous data, we first consider inference for
a univariate Gaussian with unknown mean and
unknown variance. The observed data is sampled from a univariate Gaussian with
mean $\mu = 1$ and variance $\sigma^2 = 0.5^2$. Algorithm~\ref{alg:avo} is run
for 300 epochs with mini-batches of size $M=64$ and the following configuration.
For the critic, we use a 3-layer MLP with 10 hidden nodes per layer and ReLU
activations. At each epoch, Adam is run for $n_\text{critic}=100$ iterations
with a step size $\alpha=0.01$, decay rates $\beta_1=\beta_2=0.5$ and its inner
first and second moment vectors reset at each epoch in order to avoid building
momentum in staled directions.  For modeling $\mu$ and $\sigma^2$, we use
univariate Gaussian proposal distributions $q(\mu|\psi)$ and $q(\log
\sigma|\psi)$ initialized with zero mean and unit variance. At each epoch,
parameters $\psi$ are updated by making one Adam step, with $\alpha=0.01$ and
$\beta_1=\beta_2=0.1$. Penalty coefficients are set to $\lambda=0.025$ and
$\gamma=5$.

\begin{figure}
    \centering
    \vspace{-0.8cm}
    \includegraphics[width=0.48\textwidth]{figures/gaussian-gamma=5.pdf}
    \vspace{-1cm}
    \caption{Toy problem on continuous data: Gaussian with unknown mean and unknown variance.}\label{fig:gaussian}
\end{figure}
\begin{figure}
    \centering
    \vspace{-0.8cm}
    \includegraphics[width=0.48\textwidth]{figures/poisson-gamma=5.pdf}
    \vspace{-1cm}
    \caption{Toy problem on discrete data: Poisson with unknown scale.}\label{fig:poisson}
\end{figure}

@@

Similarly, we evaluate adversarial variational optimization on discrete data,
for the inference of the scale parameter $\lambda$ of a Poisson distribution.
The observed data is sampled from a univariate Poisson with scale $\lambda=7$.
We use a univariate Gaussian proposal distribution $q(\log \lambda|\psi)$
initialized with a mean at $5$ and unit variance. Hyper-parameters are otherwise
left unchanged with respect to the previous setup. Despite the discreteness
and the non-differentiability of the underlying generator, Figure~\ref{fig:poisson}
shows that inference with adversarial variational optimization works.
The upper plot in the figure shows that $p(\mathbf{x}|\psi)$ closely
matches $p_r(\mathbf{x})$, with the mass of the proposal distribution
concentrated at $\log 7 = 1.94$.
